title: "ddtx_filter_normalize_cluster_demuxlet"
author: "adapted from Roy Oelen"
date: "2023-06-15"
output: html_document
---

```{r header, include=FALSE}
############################################################################################################################
# Authors: Roy Oelen
# Name: ddtx_filter_normalize_cluster.Rmd
# Function: filter the object on doublets, normalize with SCT and cluster+UMAP
############################################################################################################################
```

```{r libraries, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
####################
# libraries        #
####################

# for the object containing meta and count data
library(Seurat)
# for plotting
library(ggplot2)

```

```{r functions, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
####################
# Functions        #
####################

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
####################
# Main Code        #
####################

# locations of the objects
seurat_objects_loc <- '/groups/umcg-weersma/tmp01/projects/ddtx/ongoing/seurat_preprocess_samples/'
seurat_object_demultiplexed_loc <- paste(seurat_objects_loc, 'ddtx_merged_demultiplexed_demuxlet.rds', sep = '')
seurat_object_clustered_loc <- paste(seurat_objects_loc, 'ddtx_merged_demultiplexed_clustered_demuxlet.rds', sep = '')

```

```{r read_object, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# read the object
ddtx <- readRDS(seurat_object_demultiplexed_loc)

```

### before filtering
```{r show_before_filter, include=TRUE, echo=TRUE}
# show the number of cells before filtering
ddtx
```

### filtering

The individuals that were sequenced, received a gut transplant. This means that there should be expression data of both the receipiant and the donor, as such the number of clusters from Souporcell, is twice the number of individuals sequenced. The only exception here, is 220504_lane10. Here there no donors. When filtering for doublets, we as such use the 2*n clusters for all lanes but that one.

```{r doublet_filter, include=FALSE}
# filter on doublets
ddtx <- ddtx[, (
  ddtx@meta.data$lane %in% c(
    '220504_lane01',
    '220504_lane02',
    '220504_lane03',
    '220504_lane04',
    '220504_lane05',
    '220504_lane06',
    '220504_lane07',
    '220504_lane08',
    '220504_lane09',
    '220504_lane10'
    ) &
  ddtx@meta.data$DROPLET.TYPE == "SNG"
  )]
```

### after filtering
```{r show_after_filter, include=TRUE, echo=TRUE}
# show the number of cells after filtering
ddtx
```

```{r normalize_rna, include=FALSE}
# do old-style normalization
ddtx <- NormalizeData(ddtx)
```

```{r normalize_sct, include=FALSE}
# do SCT normalization
ddtx <- SCTransform(ddtx)
```

```{r pca_umap_cluster, include=FALSE}
# set the seed for reproduceability
set.seed(7777)
# do PCA dimensional reduction
ddtx <- RunPCA(ddtx)
# do 2d UMAP dimensional reduction
ddtx <- RunUMAP(ddtx, dims = 1:30, return.model = T)
# create neighbourhood graph
ddtx <- FindNeighbors(ddtx, dims = 1:30, k.param = 20)
# find clusters
ddtx <- FindClusters(ddtx, resolution = 1.5)
```

```{r donors, include=FALSE}
# add preliminary donor assignment
ddtx@meta.data[['donor_final']] <- NA
# we use the donor+receipiant for all lanes
ddtx@meta.data[ddtx@meta.data$lane %in% c('220504_lane01', '220504_lane02', '220504_lane03', '220504_lane04', '220504_lane05', '220504_lane06', '220504_lane07', '220504_lane08',     '220504_lane09', '220504_lane10'), 'donor_final'] <- ddtx@meta.data[ddtx@meta.data$lane %in% c('220504_lane01', '220504_lane02', '220504_lane03', '220504_lane04', '220504_lane05', '220504_lane06', '220504_lane07', '220504_lane08', '220504_lane09', '220504_lane10'), 'SNG.BEST.GUESS']

```

```{r create_plots, include=FALSE}
# plot the clusters
p_clusters <- DimPlot(ddtx, reduction = 'umap', group.by = 'seurat_clusters', label = TRUE, label.size = 3, repel = TRUE, raster = FALSE) + NoLegend() + ggtitle('Leiden clusters')
# plot the individuals
p_individuals <- DimPlot(ddtx, reduction = 'umap', group.by = 'donor_final', label = FALSE, label.size = 3, repel = TRUE, raster = FALSE) + ggtitle('genotyped individuals')
```

### after 2d UMAP reduction and clustering

```{r show_plots, include = TRUE, fig.width=10, fig.height=10}
p_clusters
p_individuals
```

```{r save, include=FALSE}
# save the result
saveRDS(ddtx, seurat_object_clustered_loc)
```
